### 2016-11-17 capability-machine.tex
added revoked heap segment to observation relation
revised awkward example and correctness lemma

\section{Examples}
\label{sec:examples}
\subsection{Simple Examples Using Local Capabilities}
\label{subsec:example-loc-cap}
In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1. The assertion flag is assumed to start as 0 and is different from function to function (so in the program below, \texttt{adv} cannot directly use the assertion flag of \texttt{f}). Notice that an assertion will never terminate in a $\failed$ state.
\subsubsection{Simple use of local capabilities}
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv(l);
          l := 1;
          adv(0);
          assert(l == 1)
\end{verbatim}
In this example \texttt{let l = 1 in} allocates a new local capability \texttt{l} with read-write permissions. Assuming \texttt{adv} has no access to capabilities with permit write local, they cannot store \texttt{l} and thus change its value in the second call.
\lars{So this means that the value returned by *l is 1, correct ? We
  could add an assert statement to the example to show it more directly.}
\lau{Yes, and done.}

@@

\subsection{Simple well-bracketedness/local capability example}
\begin{verbatim}
let f = fun adv =>
          let x = 1 in
          adv();
          assert(x == 1);
          x := 2;
          adv();
          return
\end{verbatim}
A sensible calling scheme uses local capabilities for the continuation capability passed to the callee. In the above example, if \texttt{adv} can store the continuation, then 
\texttt{adv} can invoke it during the second call and thus force the assertion to fail. We expect this program to either not terminate or terminate in a failed state (not caused by the assertion) or terminate with the result 0.
\lars{don't you want to allocate the location \texttt{x} first and then
  return a function taking \texttt{adv} as an argument ??}
\lau{I am not sure I get what you have in mind, could you perhaps write the program you had in mind?}

\subsection{Another simple well-bracketedness/local capability example}
Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability (but \texttt{let x = ...} produces a global capability). 
\begin{verbatim}
let f = fun adv =>
          let x = 1 in
          let l = 0 in
          adv(l);
          assert(x == 1);
          x := 2;
          adv(0);
          return
\end{verbatim}
In the above example, we give \texttt{adv} a ``mini stack'' to work with in the form of \texttt{l}. \texttt{adv} can thus store local capabilities, but their access will be revoked, when they return from the first call.

\subsection{Variant of the ``awkward'' example}
Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
\begin{verbatim}
let h = fun adv =>
          let x = 1 in
          let g = fun g' =>
                    x := 1;
                    adv(g');
                    x := 0;
                    adv();
                    assert(x == 0) in
          g(g);
          return
\end{verbatim}
In Dreyer's example, the issue arises as the context can provide the result of evaluating $\mathtt{e_1}$ (the equivalent here is \texttt{g})) to \texttt{f} (the equivalent here is \texttt{adv}). In our system, the context cannot capture this, so in the above program, \texttt{adv} is given explicit access to \texttt{g}. The lemma we would want to prove here is that the assertion does not fail.

To show this, we will probably need public/private transitions.

@@

\begin{align*}
  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \lambda W \ldotp 
                               \{ \npair{(\reg,\hs)} \mid
                             \begin{aligned}[t]
                               & \forall \heap_f, \heap', i \leq n \ldotp \\
                               & \quad(\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
                               & \qquad \Rightarrow
                               \begin{aligned}[t]
                                 & \exists W' \futurestr W \ldotp\exists \hs' \ldotp\\
                                 & \quad \heap' = \hs' \uplus \heap_f \land \\ 
                                 & \quad \heapSat[\hs']{n-i}{W'} \}
                               \end{aligned}
                             \end{aligned}
\end{align*}

>>>>

\section{Examples}
\label{sec:examples}
\subsection{Simple Examples Using Local Capabilities}
\label{subsec:example-loc-cap}
In the below, let \texttt{assert(e)} do the following: evaluate \texttt{e} if it evaluates to true, then continue execution, otherwise terminate where a designated assertion flag is set to 1. The assertion flag is assumed to start as 0 and is different from function to function (so in the program below, \texttt{adv} cannot directly use the assertion flag of \texttt{f}). Notice that an assertion will never terminate in a $\failed$ state.
\subsubsection{Simple use of local capabilities}
\begin{verbatim}
let f = fun adv =>
          let l = 1 in
          adv(l);
          l := 1;
          adv(0);
          assert(!l == 1)
\end{verbatim}
In this example \texttt{let l = 1 in} allocates a new local capability \texttt{l} with read-write permissions. Assuming \texttt{adv} has no access to capabilities with permit write local, they cannot store \texttt{l} and thus change its value in the second call.
\lars{So this means that the value returned by *l is 1, correct ? We
  could add an assert statement to the example to show it more directly.}
\lau{Yes, and done.}

@@

\subsection{Simple well-bracketedness/local capability example}
\begin{verbatim}
let f = fun adv =>
          let x = 1 in
          adv();
          assert(!x == 1);
          x := 2;
          adv();
          return
\end{verbatim}
A sensible calling scheme uses local capabilities for the continuation capability passed to the callee. In the above example, if \texttt{adv} can store the continuation, then 
\texttt{adv} can invoke it during the second call and thus force the assertion to fail. We expect this program to either not terminate or terminate in a failed state (not caused by the assertion) or terminate with the result 0.
\lars{don't you want to allocate the location \texttt{x} first and then
  return a function taking \texttt{adv} as an argument ??}
\lau{I am not sure I get what you have in mind, could you perhaps write the program you had in mind?}

\subsection{Another simple well-bracketedness/local capability example}
Here assume that \texttt{let l = ... in} produces a local, read-write permit write local capability (but \texttt{let x = ...} produces a global capability). 
\begin{verbatim}
let f = fun adv =>
          let x = 1 in
          let l = 0 in
          adv(l);
          assert(!x == 1);
          x := 2;
          adv(0);
          return
\end{verbatim}
In the above example, we give \texttt{adv} a ``mini stack'' to work with in the form of \texttt{l}. \texttt{adv} can thus store local capabilities, but their access will be revoked, when they return from the first call.

\subsection{Variant of the ``awkward'' example}
Assembly variant of the ``awkward'' example from \citep[p.~11]{Dreyer:2010:IHS:1863543.1863566}. 
\begin{verbatim}
g = fun _ => let x = 0 in
               fun f =>
                 x:=0;
                 f();
                 x:=1;
                 f();
                 assert(x==1)
\end{verbatim}
\begin{lemma}
  \newcommand{\adv}{\var{adv}}
  \newcommand{\flag}{\var{flag}}
  For all $n \in \nats$ 
  let
  \begin{align*}
    c_{\var{adv}} & \defeq ((\rwx,\glob),\start_{\adv},\addrend_{\adv},\start_{\adv}) \\
    c_{g} & \defeq ((\entry,\glob),\start_g,\addrend_g,\start_g)
  \end{align*}
if
  \[
    (\reg_0\update{\pcreg}{c_{\adv}}
          \update{r_1}{c_g},\hs_{\adv} \uplus \hs_{\flag} \uplus \hs_{g} \uplus h_f) \step[n] (\halted,h')
  \]
where 
\begin{align*}
  &\dom(hs_{\adv}) = [\start_\adv,\addrend_\adv] \\
  &\dom(hs_g) = [\start_g,\addrend_g] \\
  &\dom(hs_\flag) = [\flag,\flag] \\
  &\heapSat[\hs_{\adv}]{n}{\iota^{\var{nwl}}_{\start_\adv,\addrend_\adv}}
\end{align*}
and $hs_\flag(\flag) = 0$ and $hs_g$ contains the code of $g$ including a $\readwrite$-capability for $\hs_\flag$, 
then
\[
  h'(\flag) = 0
\]
\end{lemma}

@@

\begin{align*}
  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \lambda W \ldotp 
                               \{ \npair{(\reg,\hs)} \mid
                             \begin{aligned}[t]
                               & \forall \heap_f, \heap', i \leq n \ldotp \\
                               & \quad(\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
                               & \qquad \Rightarrow
                               \begin{aligned}[t]
                                 & \exists W' \futurestr W \ldotp\exists \hs_r, \hs' \ldotp\\
                                 & \quad \heap' = \hs' \uplus \hs_r \uplus \heap_f \land \\ 
                                 & \quad \heapSat[\hs']{n-i}{W'} \}
                               \end{aligned}
                             \end{aligned}
\end{align*}
